---
title: "Download and Unzip ABS Census Data"
author: "Jeremy Forbes"
date: "06/03/2018"
output: html_document
---

This markdown, when run, downloads the requires Commonwealth Electoral Division community profiles for 2001 and 2006. It then unzips the downloaded files and places them in the folders to be accessed by the Census-to-Dataframe file.

Profiles are split into separate folders for ABS2001 and ABS2006, of which have folders for the Zipped and Unzipped files.

#Download Census data for each CED, 2001

CED List
```{r CED_list_2001}
CED_list_2001 <- read.csv("/Users/Jeremy/Documents/R/Data/Raw-Supplement/CED_Definitions/CED2001.csv")
CED_no_2001 <- CED_list_2001$CED
```

Loop over each CED for 2001
```{r all_2001, eval = FALSE}
url_p1 = "http://www.censusdata.abs.gov.au/CensusOutput/copsub2006.NSF/All%20docs%20by%20catNo/2001~Community%20Profile~CED"
url_p2 = "/$File/BCP_CED"
url_p3 = ".zip?OpenElement"

for (i in 1:length(CED_no_2001)) {
  CED <- toString(CED_no_2001[i])
  url_temp <- paste0(url_p1, CED, url_p2, CED, url_p3)
  dest_temp = paste0("/Users/Jeremy/Documents/R/Data/Raw/ABS2001/Zipped/CED", CED, ".zip") #set destination directory
  download.file(url_temp, dest_temp, quiet = TRUE)
}
```

Unzip abs 2001
```{r unzip_abs2001, eval = FALSE}
zip_p1 <- "/Users/Jeremy/Documents/R/Data/Raw/ABS2001/Zipped/CED"
zip_p2 <- ".zip"

for (i in 1:length(CED_no_2001)) {
  CED <- toString(CED_no_2001[i])
  zip_temp <- paste0(zip_p1, CED, zip_p2)
  unzip(zip_temp, exdir = "/Users/Jeremy/Documents/R/Data/Raw/ABS2001/Unzipped", overwrite = TRUE) #set destination directory
}
```


#Download Census data for each CED, 2006 - using 2004 boundaries


CED list 2006
```{r CED_list_2006}
CED_list_2006 <- read.csv("/Users/Jeremy/Documents/R/Data/Raw-Supplement/CED_Definitions/CED2006.csv")
CED_no_2006 <- CED_list_2006$CED

CED_list_full <- read.csv("/Users/Jeremy/Documents/R/Data/Raw-Supplement/CED_Definitions/CED2006_07.csv")
CED_no_2006_07 <- CED_list_full$CED
```


Loop over each CED for 2006
```{r all_2006, eval = FALSE}
url_p1 = "http://www.censusdata.abs.gov.au/CensusOutput/copsub2006.NSF/All%20docs%20by%20catNo/2006~Community%20Profile~CED"
url_p2 = "/$File/BCP_CED"
url_p3 = ".zip?OpenElement"

for (i in 1:length(CED_no_2006)) {
  CED <- toString(CED_no_2006[i])
  url_temp <- paste0(url_p1, CED, url_p2, CED, url_p3)
  dest_temp = paste0("/Users/Jeremy/Documents/R/Data/Raw/ABS2006/Zipped/CED2004/CED", CED, ".zip") #set destination directory
  download.file(url_temp, dest_temp, quiet = TRUE)
}
```

Unzip abs 2006
```{r unzip_abs2006, eval = FALSE}
zip_p1 <- "/Users/Jeremy/Documents/R/Data/Raw/ABS2006/Zipped/CED2004/CED"
zip_p2 <- ".zip"

for (i in 1:length(CED_no_2006)) {
  CED <- toString(CED_no_2006[i])
  zip_temp <- paste0(zip_p1, CED, zip_p2)
  unzip(zip_temp, exdir = "/Users/Jeremy/Documents/R/Data/Raw/ABS2006/Unzipped/CED2004", overwrite = TRUE) #set destination directory
}
```


#Download Census data for each CED, 2006 - using 2007 boundaries

Loop over each CED for 2006 (Boundaries '07)


```{r all_2006, eval = FALSE}
url_p1 = "http://www.censusdata.abs.gov.au/CensusOutput/copsub2006.NSF/All%20docs%20by%20catNo/2006~Community%20Profile~CED07"
url_p2 = "/$File/BCP_CED07"
url_p3 = ".zip?OpenElement"

CED_no_2006_e07 = c(0)
for (i in 100:1000) {
  CED <- toString(i)
  url_temp <- paste0(url_p1, CED, url_p2, CED, url_p3)
  dest_temp = paste0("/Users/Jeremy/Documents/R/Data/Raw/ABS2006/Zipped/CED2007/CED", CED, ".zip") #set destination directory
  if (url.exists(url_temp) == TRUE) {download.file(url_temp, dest_temp, quiet = TRUE)} else {}
}

```

Unzip
```{r unzip_abs2006, eval = FALSE}
# Get IDs
CED_no_2006_e07 <- list.files("/Users/Jeremy/Documents/R/Data/Raw/ABS2006/Zipped/CED2007") %>%
  substr(4,6)

# Unzip
zip_p1 <- "/Users/Jeremy/Documents/R/Data/Raw/ABS2006/Zipped/CED2007/CED"
zip_p2 <- ".zip"

for (i in 1:length(CED_no_2006_e07)) {
  CED <- toString(CED_no_2006_e07[i])
  zip_temp <- paste0(zip_p1, CED, zip_p2)
  unzip(zip_temp, exdir = "/Users/Jeremy/Documents/R/Data/Raw/ABS2006/Unzipped/CED2007", overwrite = TRUE) #set destination directory
}
```

Trim file names
```{r}
# Trimming file names
temp <- list.files("/Users/Jeremy/Documents/R/Data/Raw/ABS2006/Unzipped/CED2007") 

for (i in 1:length(temp)) {
  file_name <- paste0("/Users/Jeremy/Documents/R/Data/Raw/ABS2006/Unzipped/CED2007/", temp[i])
  file.rename(from = file_name,
              to = gsub(" ", "", file_name))
}

remove(temp)
```


```{r}
CED_name_2006_e07 <- list.files("/Users/Jeremy/Documents/R/Data/Raw/ABS2006/Unzipped/CED2007")

CED_list_2006_e07 <- data.frame(CED = CED_no_2006_e07, Electorate = 0, State = 0)

for (i in 1:length(CED_name_2006_e07)) {
  CED_list_2006_e07$Electorate[i] = substr(strsplit(strsplit(CED_name_2006_e07[i], 'CED')[[1]][1], 'BCP-')[[1]][2],1,nchar(strsplit(strsplit(CED_name_2006_e07[i], 'CED')[[1]][1], 'BCP-')[[1]][2]) - 1)
  CED_list_2006_e07$State[i] = substr(strsplit(strsplit(CED_name_2006_e07[i], '07' )[[1]][2], 'Com')[[1]][1], 3, nchar(strsplit(strsplit(CED_name_2006_e07[i], '07' )[[1]][2], 'Com')[[1]][1]) - 1)
}
```


