---
title: "Exploring the 2011 Census data"
author: "Nicholas Tierney and Rob J Hyndman"
date: "`r Sys.Date()`"
output: 
    rmarkdown::html_vignette:
        toc: true
vignette: >
  %\VignetteIndexEntry{Exploring the 2011 Census data}
  %\VignetteEngine{knitr::rmarkdown}
  %\usepackage[utf8]{inputenc}
---


```{r setup, echo=FALSE}

knitr::opts_chunk$set(fig.width = 6,
                      fig.height = 4,
                      fig.align='center',
                      dev = "png")

```

# Introduction

`eechidna` (Exploring Election and Census Highly Informative Data Nationally for Australia) is an R package that makes it easy to look at the data from the 2011 Australian Census, and the 2013 Federal Election.

This vignette documents how to access the data from the 2011 Census. We show a few typical methods to explore the data.

## 2011 Census data

The data is loaded as `abs2016` when you load `eechidna`. Let's look take a brief glimpse of the data.

```{r load-read, message=FALSE, error = FALSE}
library(eechidna)
library(plyr)
library(dplyr)

glimpse(abs2016)

```

Here we see that we have 150 observations and 86 variables.

Each observation is data pertaining to a particular federal electorate as described by http://www.aec.gov.au/profiles/.

Each column is now described here:

```{r make-data-dictionary, include = FALSE}

data_dictionary <- tibble(
  Variable = names(abs2016), 
  Details = c(
    "Commonwealth Electoral District identifier",
    "Name of electorate",
    "State containing electorate",
    "Total population of electorate",
    "Area of electorate division in square kilometres",
    "Percentage of people aged 0-4.",
    "Percentage of people aged 5-9.",
    "Percentage of people aged 15-19.",
    "Percentage of people aged 20-24.",
    "Percentage of people aged 25-34.",
    "Percentage of people aged 35-44.",
    "Percentage of people aged 45-54.",
    "Percentage of people aged 55-64.",
    "Percentage of people aged 65-74.",
    "Percentage of people aged 75-84.",
    "Percentage of people aged 85 or higher.",
    "Percentage of people who are Australian Citizens",
    "Percentage of people who are Indigenous",
    "Percentage of people who speak a language other than English at home",
    "Percentage of people who speak only English",
    "Average number of people in a household",
    "Median weekly household income (in $)",
    "Median weekly family income (in $)",
    "Median weekly rental payment amount (of those who rent, in $)",
    "Median mortgage loan repayment amount (of mortgage payments, in $)",
    "Median age",
    "Median weekly personal income (in $)",
    "Percentage of people who are married",
    "Percentage of people who are in a de facto marriage",
    "Percentage of people who did not answer the question relating to birthplace",
    "Percentage of people born in the United Kingdom",
    "Percentage of people born in the Middle East",
    "Percentage of people born in South Eastern Europe",
    "Percentage of people born in Asia",
    "Rate of nonresponse for questions relating to language spoken at home",
    "Percentage of people affiliated with the Christian religion (of all denominations)",
    "Percentage of people affiliated with the Anglican denomimation",
    "Percentage of people affiliated with the Catholic denomimation",
    "Percentage of people affiliated with the Buddhist religion",
    "Percentage of people affiliated with the Islamic religion",
    "Percentage of people affiliated with the Jewish religion",
    "Percentage of people affiliated with a denomination of the Christian religion other than Anglican or Catholic",
    "Percentage of people with no religion",
    "Rate of nonresponse for questions relating to religion",
    "Percentage of people affiliated with a religion other than Christianity, Buddhism, Islam and Judaism",
    "Percentage of people who are currently studying",
    "Percentage of people who have completed high school",
    "Rate of nonresponse for questions relating to high school completion",
    "Rate of nonresponse for questions relating to personal income",
    "Percentage of people who work as a volunteer",
    "Rate of nonresponse for questions relating to working as a volunteer",
    "Average number of people per family",
    "Percentage of households made up of a couple with no children",
    "Percentage of households made up of a couple with children",
    "Percentage of households made up of one parent with children",
    "Percentage of people who did not answer the question relating to family income",
    "Percentage of people who did not answer the question relating to houshold income",
    "Percentage of households occupied by a single person",
    "Rate of nonresponse for questions relating to tenure",
    "Percentage of dwellings that are owned outright",
    "Percentage of dwellings that are on a mortgage",
    "Percentage of dwellings that are being rented",
    "Percentage of dwellings that are owned by the government, and rented out to tenants",
    "Rate of nonresponse for questions relating to rental costs",
    "Percentage of people with access to the internet",
    "Rate of nonresponse for questions relating to internal access",
    "Unemployment rate",
    "Labor force participation rate",
    "Percentage of people who live at a different address to what they did 5 years ago",
    "Percentage of people who have completed a Bachelor degree",
    "Percentage of people who have completed a postgraduate degree",
    "Percentage of people who have completed a diploma or certificate",
    "Rate of nonresponse for questions relating to University",
    "Percentage of employed persons who work in extractive industries (includes mining, gas, water, agriculture, waste, electricity)",
    "Percentage of employed persons who work in construction or manufacturing related industries",
    "Percentage of employed persons who work in wholesale trade, retail trade, transport, post or warehousing related industries",
    "Percentage of employed persons who work in finance or insurance related industries",
    "Percentage of employed persons who work in education and training, healthcare, wocial work, community, arts and recreation",
    "Percentage of employed persons who work in management, administration, clerical duties and sales",
    "Percentage of employed persons who work as a professional",
    "Percentage of employed persons who specialise in a trade",
    "Percentage of employed persons who work as a laborer",
    "Percentage of people who receive emuneration outside of Australia, out of the total population plus overseas visitors",
    "Percentage of people who used interent in the last week [2001 only]",
    "Rate of nonresponse for questions relating to internet use [2001 only]",
    "Percentage of people who were born overseas")
    ) # close data_frame
```

```{r display-data-dictionary, echo = FALSE}
library(knitr)
kable(data_dictionary)

```

So let's just look at some nice and simple plots using `ggplot2`.

## Unemployment 

```{r}

library(ggplot2)

ggplot(data = abs2016,
       aes(x = Unemployed)) + 
  geom_density(fill = "salmon", 
               bw = "SJ",
               colour = NA) + 
  geom_rug(colour = "salmon") +
  theme_minimal() +
  xlim(0, 12)

```


## Unemployment by state

```{r}

ggplot(data = abs2016,
       aes(x = reorder(State, -Unemployed),
           y = Unemployed,
           colour = State)) + 
  geom_boxplot() + 
  labs(x = "State",
       y = "% Unemployment") + 
  theme_minimal() + 
  theme(legend.position = "none") 
  
           
```

## Age


```{r}

ggplot(data = abs2016,
       aes(x = Age00_04)) +
   geom_density(fill = "steelblue",
               bw = "SJ",
               colour = NA) + 
  xlim(3,11) +
  geom_rug(colour = "steelblue") + 
  theme_minimal() +
  labs(x = "% Aged between 0 and 4")

```


```{r}

ggplot(data = abs2016,
       aes(x = reorder(State, -Age00_04),
           y = Age00_04,
           colour = State)) +
  geom_boxplot() + 
  theme_minimal() +
  labs(x = "State",
       y = "% Aged between 0 and 4") +
  theme(legend.position = "none") + 
  coord_flip()

```

However, there are many age groups. To look at all of them at once, we can gather them into a dataframe ready for plotting using `tidyr`.

```{r}

library(tidyr)

abs2016 %>%
  select(starts_with("Age"), 
         DivisionNm) %>%
  gather(key = "Age",
         value = "Percent_in_electorate",
         -DivisionNm) %>% 
  ggplot(data = .,
         aes(x = reorder(Age, - Percent_in_electorate),
             y = Percent_in_electorate,
             colour = Age)) +
  geom_boxplot() + 
  coord_flip() + 
  theme_minimal() + 
  theme(legend.position = "none") +
  labs(x = "Age Groups",
       y = "% in Electorate")


```

## Personal Income

```{r}

ggplot(data = abs2016,
       aes(x = MedianPersonalIncome)) + 
  geom_density(fill = "salmon",
               bw = "SJ",
               colour = NA) + 
  xlim(250,1100) +
  geom_rug(colour = "salmon") + 
  theme_minimal()
  
```

## Income by State

```{r}

ggplot(data = abs2016,
       aes(x = reorder(State, -MedianPersonalIncome),
           y = MedianPersonalIncome,
           colour = State)) + 
  geom_boxplot() + 
  theme_minimal() + 
  theme(legend.position = "none") + 
  labs(x = "State")

```

If you're intersted in getting a sense of the distribution of the data, you can add in the points to get a bit more of a sense on the distribution.

```{r}

ggplot(data = abs2016,
       aes(x = reorder(State, -MedianPersonalIncome),
           y = MedianPersonalIncome,
           colour = State)) + 
  geom_boxplot() + 
  geom_jitter(alpha = 0.35, 
              size = 2,
              width = 0.3) +
  theme_minimal() + 
  theme(legend.position = "none") + 
  labs(x = "State")

```

## Education

##### Bachelor

```{r}

ggplot(data = abs2016,
       aes(x = Bachelor)) +
  geom_density(fill = "salmon",
               bw = "SJ",
               colour = NA) + 
  geom_rug(colour = "salmon") + 
  theme_minimal() + 
  labs(x = "% of electorate with a Bachelor degree") +
  xlim(0, 30)
           

```

##### Bachelor by state

```{r}

ggplot(data = abs2016,
       aes(x = reorder(State, -Bachelor),
           y = Bachelor,
           colour = State)) +
  geom_boxplot() +
  theme_minimal() +
  labs(x = "State") + 
  theme(legend.position = "none")

```

##### Bachelor and income

```{r}

ggplot(data = abs2016,
       aes(x = Bachelor,
           y = MedianPersonalIncome)) + 
  geom_point(colour = "steelblue",
             alpha = 0.75) + 
  theme_minimal()

```


## Postgraduate


```{r}

ggplot(data = abs2016,
       aes(x = reorder(State, -Postgraduate),
           y = Postgraduate,
           colour = State)) +
  geom_boxplot() +
  theme_minimal() +
  labs(x = "State") + 
  theme(legend.position = "none")

```

##### Postgraduate and income

```{r}

ggplot(data = abs2016,
       aes(x = Postgraduate,
           y = MedianPersonalIncome)) + 
  geom_point(colour = "steelblue",
             alpha = 0.75) + 
  theme_minimal()
  
```

##### Comparing income across Bachelors and postgraduate

```{r}

abs2016 %>%
  select(Postgraduate,
         Bachelor,
         MedianPersonalIncome) %>% 
  gather(key = "Education",
         value = "Prop_Educated",
         -MedianPersonalIncome) %>%
ggplot(data = ,
       aes(x = Prop_Educated,
           y = MedianPersonalIncome,
           colour = Education)) + 
  geom_point() + 
  geom_smooth() +
  theme_minimal() +
  scale_color_brewer(type = "qual", palette = "Set1")
  # theme(legend.position = "bottom",
  #       legend.direction = "vertical")

```

## Religion

Let's look at all of the religions

```{r}

abs2016 %>%
  select(Christianity,
         Catholic,
         Buddhism,
         Islam,
         Judaism,
         NoReligion) %>%
  gather(key = "ReligionType",
         value = "Percent") %>%
  ggplot(data = .,
         aes(x = reorder(ReligionType, -Percent),
             y = Percent,
             colour = ReligionType)) + 
  geom_boxplot() + 
  theme_minimal() + 
  theme(legend.position = "none") +
  coord_flip() + 
  labs(x = "Religion")
             

```

##### Christianity by State


```{r}

ggplot(data = abs2016,
       aes(x = reorder(State, -Christianity),
           y = Christianity,
           colour = State)) + 
  geom_boxplot() +
  theme_minimal() +
  theme(legend.position = "none") +
  coord_flip() + 
  labs(x = "State")
  
           

```

## Internet access

```{r}

ggplot(data = abs2016,
       aes(x = InternetAccess)) +
  geom_density(fill = "steelblue",
               bw = "SJ",
               colour = NA) + 
  geom_rug(colour = "steelblue") + 
  theme_minimal() + 
  labs(x = "% of electorate with Internet") +
  xlim(85, 100)

```


## Internet access by state

```{r}

ggplot(data = abs2016,
       aes(x = reorder(State, -InternetAccess),
           y = InternetAccess,
           colour = State)) + 
  geom_boxplot() +
  theme_minimal() +
  theme(legend.position = "none") +
  coord_flip() + 
  labs(x = "State")

```


