---
title: "Using echidnaR"
author: "Di Cook, Heike Hofmann, Rob Hyndman, Ben Marwick, Carson Sievert, Nicholas Tierney, Fang Zhou."
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


```{r setup, echo=FALSE}

knitr::opts_chunk$set(fig.width = 6,
                      fig.height = 4,
                      fig.align='center',
                      dev = "png")

```


```{r, echo=FALSE}

# load_all(pkg = "echidnaR/")

```

`echidnaR` (Electoral Commission Handling Intelligent Data Nationally for Australia using R) is an R package that makes it easy to look at the data from the 2011 Australian Census, and the 2013 Federal Election.

This vignette documents how to access the 2011 Census data, and what it contains.

The data is loaded as `abs2011`. Let's look take a brief glimpse of the data.

```{r load-read, message=FALSE, error = FALSE}
library(echidnaR)
library(dplyr)

glimpse(abs2011)

```

Here we see that we have 150 observations and 28 variables.

Each observation is data pertaining to a particular federal electorate as described by http://www.aec.gov.au/profiles/.

Each column is now described here:

```{r make-data-dictionary, include = FALSE}

data_dictionary <- data_frame(
  Variable = c(
    "ID",
    "Name",
    "State",
    "Population",
    "Area",
    "MedianIncome",
    "Unemployed",
    "Bachelor",
    "Postgraduate",
    "Christianity",
    "Catholic",
    "Buddhism",
    "Islam",
    "Judaism",
    "NoReligion",
    "Age00_04",
    "Age05_14",
    "Age15_19",
    "Age20_24",
    "Age25_34",
    "Age35_44",
    "Age45_54",
    "Age55_64",
    "Age65_74",
    "Age75_84",
    "Age85plus",
    "BornOverseas",
    "Indigenous",
    "EnglishOnly",
    "OtherLanguageHome",
    "Married",
    "DeFacto",
    "FamilyRatio",
    "Internet", 
    "NotOwned"
    ), 
  Details = c(
    "Commonwealth Electoral District identifier",
    "Name of electorate",
    "State containing electorate",
    "Total population of electorate",
    "Area of electorate in square kilometres",
    "Median income of people within electorate",
    "Percentage of people unemployed",
    "Percentage of people whose highest qualification is a Bachelor degree",
    "Percentage of people whose highest qualification is a postgraduate degree",
    "Percentage of people affiliated with the Christian religion (of all denominations)",
    "Percentage of people affiliated with the Catholic denomimation.",
    "Percentage of people affiliated with the Buddhist religion.",
    "Percentage of people affiliated with the Islam religion.",
    "Percentage of people affiliated with the Jewish religion. ",
    "Percentage of people with no religion.",
    "Percentage of people aged 0-4.",
    "Percentage of people aged 5-9.",
    "Percentage of people aged 15-19.",
    "Percentage of people aged 20-24.",
    "Percentage of people aged 25-34.",
    "Percentage of people aged 35-44.",
    "Percentage of people aged 45-54.",
    "Percentage of people aged 55-64.",
    "Percentage of people aged 65-74.",
    "Percentage of people aged 75-84.",
    "Percentage of people aged 85 or higher.",
    "Percentage of people born outside Australia.",
    "Percentage of people who are Indigenous",
    "Percentage of people who speak only English",
    "Percentage of people who speak a language other than English at home",
    "Percentage of people who are married",
    "Percentage of people who are in a de facto marriage",
    "Total number of families to total number of people (times 100)",
    "Percentage of people with home internet",
    "Percentage of dwellings not owned (either outright or with a mortgage)")
) # close data_frame
```

```{r display-data-dictionary, echo = FALSE}
library(knitr)
kable(data_dictionary)

```

# General usage

So let's just look at some nice and simple plots.

## Age


```{r}

library(tidyr)
library(ggplot2)

abs2011 %>%
  select(starts_with("Age"), Name) %>%
  gather(key = "Age",
         value = "Percent_in_electorate",
         -Name) %>% 
  ggplot(data = .,
         aes(x = Age,
             y = Percent_in_electorate,
             colour = Age)) +
  geom_boxplot() + 
  coord_flip() + 
  theme_minimal() + 
  theme(legend.position = "none") +
  labs(y = "% in Electorate")


```

## Unemployment 

```{r}

ggplot(data = abs2011,
       aes(x = Unemployed)) + 
  geom_density(fill = "salmon", 
               bw = "SJ",
               colour = NA) + 
  geom_rug(colour = "salmon") +
  theme_minimal() +
  xlim(0, 12)

```


## Unemployment by state

```{r}

ggplot(data = abs2011,
       aes(x = reorder(State, -Unemployed),
           y = Unemployed,
           colour = State)) + 
  geom_boxplot() + 
  labs(x = "State",
       y = "% Unemployment") + 
  theme_minimal() + 
  theme(legend.position = "none") 
  
           
```


## Income

```{r}

ggplot(data = abs2011,
       aes(x = MedianIncome)) + 
  geom_density(fill = "salmon",
               bw = "SJ",
               colour = NA) + 
  geom_rug(colour = "salmon") + 
  theme_minimal()
  
```

## Income by State

```{r}

ggplot(data = abs2011,
       aes(x = reorder(State, -MedianIncome),
           y = MedianIncome,
           colour = State)) + 
  geom_boxplot() + 
  theme_minimal() + 
  theme(legend.position = "none") + 
  labs(x = "State")

```

Just a thought, you can add in the points to get a bit more information on the distribution

```{r}

ggplot(data = abs2011,
       aes(x = reorder(State, -MedianIncome),
           y = MedianIncome,
           colour = State)) + 
  geom_boxplot() + 
  geom_jitter(alpha = 0.35, 
              size = 0.5,
              width = 0.3) +
  theme_minimal() + 
  theme(legend.position = "none") + 
  labs(x = "State")

```

## Education

### Bachelor

```{r}

ggplot(data = abs2011,
       aes(x = Bachelor)) +
  geom_density(fill = "salmon",
               bw = "SJ",
               colour = NA) + 
  geom_rug(colour = "salmon") + 
  theme_minimal() + 
  labs(x = "% of electorate with a Bachelor degree") +
  xlim(0, 30)
           

```

### Bachelor by state

```{r}

ggplot(data = abs2011,
       aes(x = reorder(State, -Bachelor),
           y = Bachelor,
           colour = State)) +
  geom_boxplot() +
  theme_minimal() +
  labs(x = "State") + 
  theme(legend.position = "none")

```

### Bachelor and income

```{r}

ggplot(data = abs2011,
       aes(x = Bachelor,
           y = MedianIncome)) + 
  geom_point(colour = "steelblue",
             alpha = 0.75) + 
  theme_minimal()

```


## Postgraduate


```{r}

ggplot(data = abs2011,
       aes(x = reorder(State, -Postgraduate),
           y = Postgraduate,
           colour = State)) +
  geom_boxplot() +
  theme_minimal() +
  labs(x = "State") + 
  theme(legend.position = "none")

```

### Postgraduate and income

```{r}

ggplot(data = abs2011,
       aes(x = Postgraduate,
           y = MedianIncome)) + 
  geom_point(colour = "steelblue",
             alpha = 0.75) + 
  theme_minimal()
  
```

### Comparing income across Bachelors and postgraduate

```{r}

abs2011 %>%
  select(Postgraduate,
         Bachelor,
         MedianIncome) %>% 
  gather(key = "Education",
         value = "Prop_Educated",
         -MedianIncome) %>%
ggplot(data = ,
       aes(x = Prop_Educated,
           y = MedianIncome,
           colour = Education)) + 
  geom_point() + 
  geom_smooth() +
  theme_minimal() +
  scale_color_brewer(type = "qual", palette = "Set1")
  # theme(legend.position = "bottom",
  #       legend.direction = "vertical")

```

## Religion

Let's look at al of the religions

```{r}

abs2011 %>%
  select(Christianity,
         Catholic,
         Buddhism,
         Islam,
         Judaism,
         NoReligion) %>%
  gather(key = "ReligionType",
         value = "Percent") %>%
  ggplot(data = .,
         aes(x = reorder(ReligionType, -Percent),
             y = Percent,
             colour = ReligionType)) + 
  geom_boxplot() + 
  theme_minimal() + 
  theme(legend.position = "none") +
  coord_flip() + 
  labs(x = "Religion")
             

```

### Christianity by State


```{r}

ggplot(data = abs2011,
       aes(x = reorder(State, -Christianity),
           y = Christianity,
           colour = State)) + 
  geom_boxplot() +
  theme_minimal() +
  theme(legend.position = "none") +
  coord_flip() + 
  labs(x = "State")
  
           

```

## Internet

```{r}

ggplot(data = abs2011,
       aes(x = Internet)) +
  geom_density(fill = "steelblue",
               bw = "SJ",
               colour = NA) + 
  geom_rug(colour = "steelblue") + 
  theme_minimal() + 
  labs(x = "% of electorate with Internet") +
  xlim(85, 100)

```


## Internet by state

```{r}

ggplot(data = abs2011,
       aes(x = reorder(State, -Internet),
           y = Internet,
           colour = State)) + 
  geom_boxplot() +
  theme_minimal() +
  theme(legend.position = "none") +
  coord_flip() + 
  labs(x = "State")

```


# Spatial Mapping

One of the really amazing things about this R package is that it contains the shapefiles of the electorates. These are coming soon!

Note to writers:

- It would be really great to look at visualisations that show **where do ... live**.
