---
title: "Using echidnaR: 2011 Census data"
author: "Di Cook, Heike Hofmann, Rob Hyndman, Ben Marwick, Carson Sievert, Nicholas Tierney, Fang Zhou."
date: "`r Sys.Date()`"
output: 
    rmarkdown::html_vignette:
        toc: true

vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


```{r setup, echo=FALSE}

knitr::opts_chunk$set(fig.width = 6,
                      fig.height = 4,
                      fig.align='center',
                      dev = "png")

```


```{r, echo=FALSE}

# load_all(pkg = "echidnaR/")

```

# Introduction

`echidnaR` (Electoral Commission Handling Intelligent Data Nationally for Australia using R) is an R package that makes it easy to look at the data from the 2011 Australian Census, and the 2013 Federal Election.

This vignette documents how to access the data from the 2013 Federal Election (House of Representatives (first preferences only). In this vignette we show a few typical methods to explore the data.

## 2013 Federal Election House of Representatives data

In the `echidnaR` package the election dataset is called `aec2013`, and can be inspected like this:

```{r load-read, message=FALSE, error = FALSE}
library(echidnaR)
library(dplyr)

glimpse(aec2013)

```


We can see that the data contains `r nrow(aec2013)` rows and `r ncol(aec2013)` columns. 

Each row represents one canditate at one polling place. We obtained the data from http://results.aec.gov.au/ . This package incorporates data that is
© Commonwealth of Australia (Australian Electoral Commission) 2016.

The table below describes the columns in the election data:

```{r make-data-dictionary, include = FALSE}

data_dictionary <- data_frame(
  Variable = c(
 "StateAb",            "DivisionID.x",       "DivisionNm.x",      
 "PollingPlaceID",     "PollingPlace",       "CandidateID",       
 "Surname",            "GivenNm",            "BallotPosition",    
 "Elected",            "HistoricElected",    "PartyAb",           
 "PartyNm",            "OrdinaryVotes",      "Swing",             
 "State",              "DivisionNm.y",       "PollingPlaceTypeID",
 "PremisesNm",         "PremisesAddress1",   "PremisesAddress2",  
 "PremisesAddress3",   "PremisesSuburb",     "PremisesStateAb",   
 "PremisesPostCode",   "Latitude",           "Longitude",         
 "uid"   
  ), 
  Details = c(
    "Commonwealth Electoral District identifier",
    'xxx')
    
  ) # close data_frame
```


```{r display-data-dictionary, echo = FALSE}
library(knitr)
kable(data_dictionary)

```

We can summarise the data to discover some basic details about the election.

```{r total_votes_for_parties}

aec2013 %>% 
  select(PartyNm, OrdinaryVotes) %>% 
  group_by(PartyNm) %>% 
  summarise(total_votes = sum(OrdinaryVotes)) %>% 
  ungroup() %>%
  arrange(desc(total_votes))
```


