---
title: "Using echidnaR: 2011 Census data"
author: "Di Cook, Heike Hofmann, Rob Hyndman, Ben Marwick, Carson Sievert, Nicholas Tierney, Fang Zhou."
date: "`r Sys.Date()`"
output: 
    rmarkdown::html_vignette:
        toc: true

vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


```{r setup, echo=FALSE}

knitr::opts_chunk$set(fig.width = 6,
                      fig.height = 4,
                      fig.align='center',
                      dev = "png")

```


```{r, echo=FALSE}

# load_all(pkg = "echidnaR/")

```

# Introduction

`echidnaR` (Electoral Commission Handling Intelligent Data Nationally for Australia using R) is an R package that makes it easy to look at the data from the 2011 Australian Census, and the 2013 Federal Election.

This vignette documents how to access the data from the 2013 Federal Election (House of Representatives (first preferences only). In this vignette we show a few typical methods to explore the data.

## 2013 Federal Election House of Representatives data

In the `echidnaR` package we have two election datasets. One is called `aec2013`, and each row represents a particular canditate at a particular polling place. We can and can inspect this data like this:

```{r load-read, message=FALSE, error = FALSE}
library(echidnaR)
library(dplyr)
# inspect the data frame
glimpse(aec2013)

```


We can see that the data contains `r nrow(aec2013)` rows and `r ncol(aec2013)` columns.

The other dataset is `aec2013_electorate` and is the results of the election aggregated for each electorate. Each row represents a particular candidate at a particular electorate.

 We obtained these data from http://results.aec.gov.au/ . This package incorporates data that is © Commonwealth of Australia (Australian Electoral Commission) 2016.

<!-- The table below describes the columns in the `aec2013` data: -->

```{r make-data-dictionary, include = FALSE}

# data_dictionary_1 <- data_frame(
#   Variable = c(
#  "StateAb",            "DivisionID.x",       "DivisionNm.x",      
#  "PollingPlaceID",     "PollingPlace",       "CandidateID",       
#  "Surname",            "GivenNm",            "BallotPosition",    
#  "Elected",            "HistoricElected",    "PartyAb",           
#  "PartyNm",            "OrdinaryVotes",      "Swing",             
#  "State",              "DivisionNm.y",       "PollingPlaceTypeID",
#  "PremisesNm",         "PremisesAddress1",   "PremisesAddress2",  
#  "PremisesAddress3",   "PremisesSuburb",     "PremisesStateAb",   
#  "PremisesPostCode",   "Latitude",           "Longitude",         
#  "uid"   
#   ), 
#   Details = c(
#     "Commonwealth Electoral District identifier",
#     'xxx')
#     
#   ) # close data_frame
```


```{r display-data-dictionary, echo = FALSE}
# library(knitr)
# kable(data_dictionary_1)

```

We can summarise the data to discover some basic details about the election. Here we can find the total number of ordinary votes (not including postal votes, etc.) for each party:

```{r total_votes_for_parties}
total_votes_for_parties <- aec2013 %>% 
  select(PartyNm, OrdinaryVotes) %>% 
  group_by(PartyNm) %>% 
  summarise(total_votes = sum(OrdinaryVotes)) %>% 
  ungroup() %>%
  arrange(desc(total_votes))
# inspect 
total_votes_for_parties
```

And we can plot this like so:

```{r plot_total_votes_for_parties,  fig.height = 6}
library(ggplot2)
library(scales)
ggplot(total_votes_for_parties, aes(reorder(PartyNm, total_votes), total_votes )) +
  geom_point() + 
  coord_flip() + 
  scale_y_continuous(labels = comma) +
  theme_bw() +
  ylab("Total ordinary votes") +
  xlab("Party") +
  theme(text = element_text(size=10))
```

We can look at the proportion of voters for a particular party in each electorate, grouped by each state. This plot was inspired by [Peter Ellis's](https://peterellisblog.com/) work on the NZ election data. 


```{r plot_prop_particular_party}
# summarise and compute proportion of votes for a particular party
p <- aec2013 %>%
  group_by(Electorate, State) %>%
  summarise(
    TotalVotes = sum(OrdinaryVotes),
    ProportionLabor = round(sum(OrdinaryVotes[PartyNm == "Australian Labor Party"]) / TotalVotes, 3)) %>%
  filter(TotalVotes != 0) %>% 
  arrange(desc(ProportionLabor)) %>% 
 group_by(State) %>% 
  # send the data to the plotting function
  do(plots=ggplot(data = .) + 
       aes(x = ProportionLabor, y = reorder(Electorate, ProportionLabor), size = TotalVotes, label = State) +
       geom_point() +
       ylab("Electorate") +
       labs(title = .$State) + 
       scale_x_continuous("Proportion voting Labor Party", label = percent) +
       scale_size("Number of\nvotes cast", label = comma)  +
       theme_bw() +
       theme(text = element_text(size=10)))

# draw the plots
library(gridExtra)
n <- length(p$plots)
nCol <- floor(sqrt(n))
do.call("grid.arrange", c(p$plots, ncol=nCol))



```


