---
title: "Using echidnaR"
author: "Di Cook, Heike Hofmann, Rob Hyndman, Ben Marwick, Carson Sievert, Nicholas Tierney, Fang Zhou."
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


```{r}

# load_all(pkg = "echidnaR/")

```

`echidnaR` is an R package that makes it easy to look at the data from the 2011 Australian Census.

This vignette (so far) documents how to access the data, and what it is.

The data is loaded as `abs2011`. Let's look take a brief glimpse of the data.

```{r load-read, message=FALSE, error = FALSE}
library(echidnaR)
library(dplyr)

glimpse(abs2011)

```

Here we see that we have 150 observations and 28 variables.

Each observation is data pertaining to a particular federal electorate as described by <url to the gov website of electorates>.

Each column is now described here:

```{r make-data-dictionary, include = FALSE}

data_dictionary <- data_frame(
  Variable = c(
    "ID",
    "Name",
    "State",
    "Population",
    "Area",
    "MedianIncome",
    "Unemployed",
    "Bachelor",
    "Postgraduate",
    "NoReligion",
    "Age0_4",
    "Age5_14",
    "Age15_19",
    "Age20_24",
    "Age25_34",
    "Age35_44",
    "Age45_54",
    "Age55_64",
    "Age65_74",
    "Age75_84",
    "Age85plus",
    "BornOverseas",
    "Indigenous",
    "EnglishOnly",
    "OtherLanguageHome",
    "Married",
    "DeFacto",
    "FamilyRatio"
    ), 
  Details = c(
    "Commonwealth Electoral District identifier",
    "Name of electorate",
    "State containing electorate",
    "Total population of electorate",
    "Area of electorate in square kilometres",
    "Median income of people within electorate",
    "Percentage of people unemployed",
    "Percentage of people whose highest qualification is a Bachelor degree",
    "Percentage of people whose highest qualification is a postgraduate degree",
    "Percentage of people with no religion.",
    "Percentage of people aged 0-4.",
    "Percentage of people aged 5-9.",
    "Percentage of people aged 15-19.",
    "Percentage of people aged 20-24.",
    "Percentage of people aged 25-34.",
    "Percentage of people aged 35-44.",
    "Percentage of people aged 45-54.",
    "Percentage of people aged 55-64.",
    "Percentage of people aged 65-74.",
    "Percentage of people aged 75-84.",
    "Percentage of people aged 85 or higher.",
    "Percentage of people born outside Australia.",
    "Percentage of people who are Indigenous",
    "Percentage of people who speak only English",
    "Percentage of people who speak a language other than English at home",
    "Percentage of people who are married",
    "Percentage of people who are in a de facto marriage",
    "Total number of families to total number of people (times 100).")
) # close data_frame
```

```{r display-data-dictionary, echo = FALSE}
library(knitr)
kable(data_dictionary)

```

# General usage

So let's just look at some nice and simple plots.

Let's see what our age population is like

```{r}

library(tidyr)
library(ggplot2)

abs2011 %>%
  select(starts_with("Age"), Name) %>%
  gather(key = "Age",
         value = "Percent_in_electorate",
         -Name) %>% 
  ggplot(data = .,
         aes(x = Age,
             y = Percent_in_electorate,
             fill = Age)) +
  geom_boxplot() + 
  coord_flip() + 
  theme(legend.position = "none")
             

```

## Unemployment 

```{r}

ggplot(data = abs2011,
       aes(x = Unemployed)) + 
  geom_density(fill = "salmon") + 
  geom_rug(colour = "salmon")

```


## Unemployment by state

```{r}

ggplot(data = abs2011,
       aes(x = reorder(State, -Unemployed),
           y = Unemployed,
           fill = State)) + 
  geom_boxplot() + 
  labs(x = "State",
       y = "% Unemployment") + 
  theme_minimal() + 
  theme(legend.position = "none") 
  
           
```


## Income

```{r}

ggplot(data = abs2011,
       aes(x = MedianIncome)) + 
  geom_density(fill = "salmon") + 
  geom_rug(colour = "salmon") + 
  theme_minimal()
  
```

## Income by State

```{r}

ggplot(data = abs2011,
       aes(x = reorder(State, -MedianIncome),
           y = MedianIncome,
           colour = State)) + 
  geom_boxplot() + 
  theme_minimal() + 
  theme(legend.position = "none") + 
  labs(x = "State")

```

Just a thought, you can add in the points to get a bit more information on the distribution

```{r}

ggplot(data = abs2011,
       aes(x = reorder(State, -MedianIncome),
           y = MedianIncome,
           colour = State)) + 
  geom_boxplot() + 
  geom_jitter(alpha = 0.35, 
              size = 0.5,
              width = 0.3) +
  theme_minimal() + 
  theme(legend.position = "none") + 
  labs(x = "State")

```

## Education

### Bachelor

```{r}

ggplot(data = abs2011,
       aes(x = Bachelor)) +
  geom_density(fill = "salmon") + 
  geom_rug(colour = "salmon") + 
  theme_minimal() + 
  labs(x = "% of individuals with a Bachelor degree")
           

```

### Bachelor by state

```{r}

ggplot(data = abs2011,
       aes(x = reorder(State, -Bachelor),
           y = Bachelor,
           colour = State)) +
  geom_boxplot() +
  theme_minimal() +
  labs(x = "State") + 
  theme(legend.position = "none")

```

### Bachelor and income

```{r}

ggplot(data = abs2011,
       aes(x = Bachelor,
           y = MedianIncome)) + 
  geom_point() + 
  theme_minimal()
  
           

```


## Postgraduate


```{r}

ggplot(data = abs2011,
       aes(x = reorder(State, -Postgraduate),
           y = Postgraduate,
           colour = State)) +
  geom_boxplot() +
  theme_minimal() +
  labs(x = "State") + 
  theme(legend.position = "none")

```

### Postgraduate and income

```{r}

ggplot(data = abs2011,
       aes(x = Postgraduate,
           y = MedianIncome)) + 
  geom_point() + 
  theme_minimal()
  
```

### Comparing income across Bachelors and postgraduate

```{r}

abs2011 %>%
  select(Postgraduate,
         Bachelor,
         MedianIncome) %>% 
  gather(key = "Education",
         value = "Prop_Educated",
         -MedianIncome) %>%
ggplot(data = ,
       aes(x = Prop_Educated,
           y = MedianIncome,
           colour = Education)) + 
  geom_point() + 
  geom_smooth() +
  theme_minimal() 
  # theme(legend.position = "bottom",
  #       legend.direction = "vertical")

```


# Spatial Mapping

One of the really amazing things about this R package is that it contains the shapefiles of the electorates. These are coming soon!

Some small gripes.

- Variable names: Should they not be capitalised?
- State names: These are capitalised for ACT and NSW, but not for Qld and Vic
