---
title: "Getting Oz Electorate shapefiles into shape"
author: "Heike Hofmann, Di Cook and Jeremy Forbes"
date: "`r Sys.Date()`"
output: 
    rmarkdown::html_vignette:
        toc: true
vignette: >
  %\VignetteIndexEntry{Getting Oz Electorate shapefiles into shape}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---


```{r setup, echo=FALSE, include = FALSE}
library(knitr)
opts_chunk$set(warning = FALSE,
               message = FALSE,
               fig.width = 6,
               fig.height = 4,
               fig.align='center',
               dev = "png")
```

This vignette details the procedure used to obtain the maps of electoral boundaries for each of the Australian federal elections and Censuses. 

The Australian Electorate Commission publishes the boundaries of the electorates on their website at [http://www.aec.gov.au/Electorates/gis/gis_datadownload.htm](http://www.aec.gov.au/Electorates/gis/gis_datadownload.htm) (2010-2016). Electoral boundaries for 2001 are sourced from the Australian Government at [https://data.gov.au/](https://data.gov.au/dataset/ds-dga-0b939a62-e53e-4616-add5-77f909b58ddd/details?q=asgc%202001). The 2004 and 2007 electoral boundaries are available from the Australian Bureau of Statistics [http://www.abs.gov.au/AUSSTATS/abs@.nsf/DetailsPage/2923.0.30.0012006?OpenDocument](http://www.abs.gov.au/AUSSTATS/abs@.nsf/DetailsPage/2923.0.30.0012006?OpenDocument).

Once the files (preferably the national files) are downloaded, unzip the file (it will build a folder with a set of files). We want to read the shapes contained in the `shp`, `TAB`, or `MIF` file into R. The `rgdal` library can be used to do this.

The function `get_electorate_shapes` in this package extracts a list from the shapefile, consisting of a `dataframe` containing coordinates of each polygon and a `dataframe` with data associated with each polygon (electorate). These can be used directly with `ggplot` graphics. Alternatively, the `load_shapefile` function (also from `eechidna`) imports the shapefile as a `SpatialPolygonsDataFrame`.

```{r, message=FALSE, warning=FALSE, eval = F}
library(eechidna)
# shapeFile contains the path to the shp file:
shapeFile <- "/PATH-ON-YOUR-COMPUTER/2021_ELB_region.shp"
map_and_data <- get_electorate_shapes(shapeFile)

nat_map <- map_and_data$map
nat_data <- map_and_data$data
```

```{r, echo = F}
library(eechidna)
nat_map22 <- nat_map_download(2022)
nat_data22 <- nat_data_download(2022)
nat_map <- nat_map22
nat_data <- nat_data22
```

Here we have a map of the Australian electorates colored by their size as given in the data (variable `area_sqkm`):

```{r, message=FALSE, warning=FALSE, fig.width=8, fig.height=6}
library(ggplot2)
library(ggthemes)
library(ggmap)

ggplot(aes(map_id=id), data=nat_data) +
  geom_map(aes(fill=area_sqkm), map=nat_map) +
  expand_limits(x=nat_map$long, y=nat_map$lat) + 
  theme_map() + coord_equal()
```

The `get_electorate_shapes` function was re-written Australian electoral boundaries for 2022, but may need some tweaking for future electoral maps. (Code for previous years can be found in release v1.4.1.) Each step of this function is detailed below, with the running example of the Australian electoral boundaries for 2022.

## Steps

For the 2022 election, the national electorate boundaries are given in ESRI `shp` format.

```{r eval=FALSE, echo=FALSE}
library(sf)

# shapeFile contains the path to the shp file:
shapeFile <- "/PATH-TO-YOUR-COMPUTER/national-midmif-09052016/COM_ELB.TAB"

sF <- read_sf(shapeFile)
```

`sF` is a spatial data frame containing all of the polygons. First, for convenience, lets change all variable names in the associated data set to lower case.

```{r, eval=FALSE, message = FALSE}
names(sF) <- tolower(names(sF))
```

We now use `rmapshaper` to thin the polygons and ensure that there are no holes while preserving the geography:

```{r eval=FALSE, message=FALSE}
library(rmapshaper)
```

```{r, eval=FALSE, message=FALSE}
sF_polys <- rmapshaper::ms_simplify(sF, keep = 0.001)
```

`keep` is the numerical value indicating proportion of vertices to keep in the map, reducing the number of points. Doing this helps reduce the overall size of the map considerably, making it faster to plot. For data analysis, you don't need detailed maps.

### Extracting the electorate information 

A spatial polygons data frame consists of both a data set with information on each of the entities (in this case, electorates), and a set of polygons for each electorate (sometimes multiple polygons are needed, e.g. if the electorate has islands). We want to extract both of these parts.

```{r eval=FALSE}
nat_data <- st_set_geometry(sF, NULL)
head(nat_data)
```

The row names of the data file are identifiers corresponding to the polygons - we want to make them a separate variable:
```{r, eval=FALSE, message=FALSE}
nat_data$id <- row.names(nat_data)
```
In the currently published version of the 2022 electorate boundaries, the `data` data frame has variable `elect_div` of the electorates' names, but not state information. We add this with:
```{r eval=FALSE}
  if (!("state" %in% names(nat_data))) {
    states <- states22
    states$elect_div <- toupper(states$elect_div)
    nat_data <- nat_data %>%
      left_join(states) %>%
      select(id, elect_div, state, numccds, area_sqkm, long_c, lat_c)
  }
```

giving the column `state`, which is an abbreviation of the state name. It might be convenient to merge this information (or at least the state abbreviation) into the polygons (see below).

We are almost ready to export this data into a file, but we still want include geographic centers in the data (see also below).

### Extracting the polygon information 

The `sptable` function in the `spbabel` package extracts the polygons into a data frame. 
```{r eval=FALSE}
  nat_map <- spbabel::sptable(sF_polys)
```

We need to make sure that `group` and `piece` are kept as factor variables - if they are allowed to be converted to numeric values, it messes things up, because as factor levels `9` and `9.0` are distinct, whereas they are not when interpreted as numbers ...
```{r eval=FALSE}
nat_map$group <- paste("g",nat_map$piece, sep=".")
```

It is useful to have the electorate name and state attached to the map.

```{r eval=FALSE}
nms <- nat_data %>% select(id, elect_div, state)
nat_map$id <- as.character(nat_map$id)
nat_map <- dplyr::left_join(nat_map, nms, by="id")
```

The map data is ready to be exported to a file:

```{r eval=FALSE}
head(nat_map)
```

### Getting centroids 

Getting centroids or any other information from a polygon is fairly simple, once you have worked your way through the polygon structure. The `sf` package makes this easier to do now. We will wrap this into a little function called `centroid` to help us with that:

```{r, eval=FALSE, message=FALSE, warning=FALSE}
library(purrr)
centroid <- function(i, polys) {
  ctr <- st_centroid(st_geometry(sF)[[i]])
  data.frame(long_c=ctr[1], lat_c=ctr[2])
}

centroids <-  purrr::map_df(1:nrow(sF), centroid, polys=sf)

head(centroids)
```

The centroids come in the same order as the data (luckily) and we just extend the data set (dropping the geometry) for the electorates by this information, and finally export:

```{r, eval=FALSE, message=FALSE, warning=FALSE}
nat_data <- st_set_geometry(sF, NULL)
nat_data <- data.frame(nat_data, centroids)
```

```{r, eval=FALSE}
write.csv(nat_data, "National-data-2022.csv", row.names=FALSE)
```

Finally, just to check the data, after running `get_electorate_shapes()`, a map of the Australian electorates colored by their size as given in the data (variable `area_sqkm`):

```{r, eval=FALSE, message=FALSE, warning=FALSE, fig.width=8, fig.height=6}
ggplot(aes(map_id=id), data=nat_data) +
  geom_map(aes(fill=area_sqkm), map=nat_map) +
  expand_limits(x=nat_map$long, y=nat_map$lat) + 
  theme_map() + coord_equal()
```

