---
title: "Mapping Australia's Electorates"
author: "Nathaniel Tomasetti, Di Cook, Heike Hofmann"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

We're interested in making a map of the latest Australian election results. To do so we need to take the electoral regions, merge with election results and plot the colored polygons. However, the Australian electoral map is not conducive to chloropleth map, because most of the regions concentrate in the five big cities: Sydney, Melbourne, Brisbane, Adelaide and Perth.  

We need to start with the `nat_map` and merge this with the election results.

```{r load-read, message=FALSE, error = FALSE, warning=FALSE, echo=FALSE}
library(eechidna)
library(dplyr)
library(ggplot2)
data(nat_data_cart)
data(nat_map)
data(aec2013_fp_electorate)
map.winners <- aec2013_fp_electorate %>% filter(Elected == "Y") %>% 
  select(Electorate, PartyNm) %>% 
  merge(nat_map, by.x="Electorate", by.y="ELECT_DIV")

#Grouping different Lib/Nats togethers
map.winners$PartyNm <- as.character(map.winners$PartyNm)
coalition <- c("Country Liberals (NT)", "Liberal", "Liberal National Party of Queensland", "The Nationals")
map.winners.grouped <- mutate(map.winners, PartyNm = ifelse(as.character(PartyNm) %in% coalition,
                                                      "Liberal National Coalition", PartyNm))
map.winners.grouped <- map.winners.grouped %>% arrange(group, order)

# Colour cells to match that parties colours
# Order = Australian Labor Party, Independent, Katters, Lib/Nats Coalition, Palmer, The Greens
partycolours = c("#FF0000", "#000000", "#550000", "#0000FF", "#FF9900", "#00FF00")

# Make it look like a map
theme_map <- theme_bw()
theme_map$line <- element_blank()
theme_map$strip.text <- element_blank()
theme_map$axis.text <- element_blank()
theme_map$plot.title <- element_blank()
theme_map$axis.title <- element_blank()
theme_map$panel.border <- element_rect(colour = "grey90", size=1, fill=NA)
```


```{r fig.width=9, fig.height=6, echo=FALSE, warning=FALSE, message=FALSE, error=FALSE}
ggplot(data=map.winners.grouped) + 
  geom_polygon(aes(x=long, y=lat, group=group, order=order, fill=PartyNm)) +
  scale_fill_manual(name="Politcal Party", values=partycolours) +
  theme_map + coord_equal() + theme(legend.position="bottom")
```

Now lets overlay some cartograms on the electorates. 

All that's left is colouring these by the current party holding each electorate.
We're going to group the entire Liberal/National Coalition together under the colour blue.

```{r fig.width=9, fig.height=6, echo=FALSE, warning=FALSE, message=FALSE, error=FALSE}
# Load election results
cart.winners <- aec2013_fp_electorate %>% filter(Elected == "Y") %>% 
  select(Electorate, PartyNm) %>% 
  merge(nat_data_cart, by.x="Electorate", by.y="ELECT_DIV")

#Grouping different Lib/Nats togethers
cart.winners$PartyNm <- as.character(cart.winners$PartyNm)
coalition <- c("Country Liberals (NT)", "Liberal", "Liberal National Party of Queensland", "The Nationals")
cart.winners.grouped <- mutate(cart.winners, PartyNm = ifelse(as.character(PartyNm) %in% coalition,
                                                      "Liberal National Coalition", PartyNm))

# Plot it
ggplot(data=nat_map) +
  geom_polygon(aes(x=long, y=lat, group=group, order=order),
               fill="grey90", colour="white") +
  geom_point(data=cart.winners.grouped, aes(x=x, y=y, colour=PartyNm), size=2, alpha=0.8) +
  scale_colour_manual(name="Politcal Party", values=partycolours) +
  theme_map + coord_equal() + theme(legend.position="bottom")

```

